---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

## Loading and preprocessing the data

Set global option

```{r setOptions}
options(scipen=1, digits=2)
require(ggplot2)
```

Load the data. 

```{r loadData}
data <- read.csv("activity/activity.csv")
str(data)
```

Transform the `date` attribute to `Date` type

```{r transformData}
data$date <- as.Date(data$date)
str(data)
```

## What is mean total number of steps taken per day?

Aggregate the data.frame `data` by `date` to get the total number of steps taken each day

```{r aggregateDayData}
dataPerDay <- aggregate(data$steps, by = list(data$date), FUN = sum)
names(dataPerDay) <- c("date", "steps")
str(dataPerDay)
```

Histogram of the total number of steps taken each day

```{r histogramPlot, fig.height=4}
ggplot(data=dataPerDay, aes(steps)) + geom_histogram()
```

Calculat the **mean** and **median** total number of steps taken per day

```{r calcMean&Median}
stepMean <- mean(dataPerDay$steps, na.rm = TRUE)
stepMedian <- median(dataPerDay$steps, na.rm = TRUE)
```

The **mean** total number of steps taken per day is `r stepMean` and the **median** total number of steps taken per day is `r stepMedian`.

## What is the average daily activity pattern?

Aggregate the data.frame `data` by `interval` to get the average number of steps taken, averaged across all days

```{r aggregateIntervalData}
dataPerInterval = aggregate(data$steps, by = list(data$interval), FUN = mean,  na.rm=TRUE, na.action=NULL)
names(dataPerInterval) <- c("interval", "steps")
str(dataPerInterval)
```

Plot the time series plot of the data.frame `dataPerInterval`

```{r timeSeriesPlot}
ggplot(data=dataPerInterval, aes(interval, steps)) + geom_line()
```

Get the interval containing the maximum number of steps

```{r getMaxInterval}
index <- which.max(dataPerInterval$steps)
intvl <- dataPerInterval$interval[[index]]
```

The max interval is `r intvl`.

## Imputing missing values

Get the total number of missing values in the dataset

```{r getNullNumber}
numNA <- sum(is.na(data))
```

The total number of missing values in the dataset is `r numNA`.

We use the for that 5-minite interval to fill in all of the missing values in the dataset.

Filling the data set `dataNonNa`
```{r fillNa}
dataNonNa <- data
for (i in 1:length(dataNonNa$steps)){
  if (is.na(dataNonNa$steps[i])){
    j = i %% length(dataPerInterval$steps)
    if (j == 0) j = length(dataPerInterval$steps)
    dataNonNa$steps[i] <- dataPerInterval$steps[j]
  }
}
str(dataNonNa)
```

Aggregate and plot the new histogram

```{r aggreHist}
dataPerDayNonNa <- aggregate(dataNonNa$steps, by = list(dataNonNa$date), FUN = sum)
names(dataPerDayNonNa) <- c("date", "steps")
str(dataPerDayNonNa)
ggplot(data=dataPerDayNonNa, aes(steps)) + geom_histogram()
```

Get the **mean** and **median** total number of steps taken per day without NAs.
```{r calcMean&MedianWithoutNAs}
stepMeanNonNA <- mean(dataPerDayNonNa$steps, na.rm = TRUE)
stepMedianNonNA <- median(dataPerDayNonNa$steps, na.rm = TRUE)
```

The **mean** total number of steps taken per day without NAs is `r stepMeanNonNA` and the **median** total number of steps taken per day is `r stepMedianNonNA`.
And the previous **mean** value is `r stepMean` and previous **meadian** value is `r stepMedian`. So we can find that the **mean** value is same because we use the the mean for that 5-minute interval to fill the NAs, and because all values of some days is all NA, so after being filled these total number of is the **mean** value so the **mean** value become the **median** value too.

## Are there differences in activity patterns between weekdays and weekends?

So we modify the `dataNonNa` for the convinience of the pattern analysis between weekdays and weekends.

```{r modifyData}
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
dataNonNa$wDay <- factor((weekdays(dataNonNa$date) %in% weekdays1), 
         levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))
str(dataNonNa)
```

Aggregate `dataNonNa` by `interval` and `wDay`

```{r aggregateDataIntervalNonNa}
dataPerIntervalNonNa <- aggregate(dataNonNa$steps, by = list(dataNonNa$interval, dataNonNa$wDay), FUN = mean)
names(dataPerIntervalNonNa) <- c("interval", "wDay", "steps")
str(dataPerIntervalNonNa)
```

Plot the panel plot

```{r ggplotWeekendOrWeekday}
ggplot(dataPerIntervalNonNa, aes(interval, steps)) + geom_line() + facet_grid(wDay ~ .) 
```

From the plot we can find that steps curve has one peak more higher than others on weekdays. Instead steps curve more uniform on weekends.

